### 2018-11-14 writing/thesis/chapters/methods.tex
Rewrite section about cloud scenarios.

% TODO: Extract to settings.tex
\newcommand{\Qrho}{\rho}
\newcommand{\Qj}{\rho \bm{v}}
%\newcommand{\Qv}{\ensuremath{\Qrho^{-1} \Qj}}
\newcommand{\Qv}{\bm{v}}
\newcommand{\QE}{\rho E}
\newcommand{\stressT}{\bm{\sigma}}
\newcommand{\pressure}{p}
\newcommand{\maxConvEigen}{\vert \lambda_c^{\text{max}} \vert}
\newcommand{\maxViscEigen}{\vert \lambda_v^{\text{max}} \vert}

@@

We use the equation of state of an ideal gas
\begin{equation}
  \pressure = (\gamma - 1) \left(\QE - 0.5 \left(\Qv \cdot \Qj \right) \right).
\end{equation}
The temperature $T$ relates to pressure and density by the ideal gas law
\begin{equation}
  \label{eq:temperature}
 \frac{\pressure}{\Qrho} = RT,
\end{equation}
where $R$ is the specific gas constant.

@@

\section{Scenarios}
\label{sec:scenarios}

\subsection{Clouds}
Air is initially at rest and in hydrostractic balance
\begin{equation}
  \label{eq:hydrostatic-balance}
 \pdv{p (z)}{z} = -\rho(z) g.
\end{equation}
Inserting \cref{eq:temperature} leads to the \textsc{ode}
\begin{equation}
  \pdv{p(z)}{z} = - \frac{p(z)}{RT} g,
\end{equation}
which can be solved by seperation of variables.
This results in the equation
\begin{equation}
  \label{eq:bubble-pressure}
  p(z) = c \exp \left( - \frac{gz}{RT} \right),
\end{equation}
where the constant of integration $c$ can be found by considering the case
\begin{equation}
  p(0) = c.
\end{equation}

The scenario is described in terms of potential temperature $\theta$
\begin{equation}
  \theta = T \frac{p_0}{p}^{R/c_p},
\end{equation}
where $p_o = \SI{10e5}{\Pa}$ is the reference pressure.
Solving for $T$ leads to
\begin{equation}
  T = \theta (\frac{p}{p_0})^{R/c_p}.
\end{equation}

Note: Above approach leads to wrong atmosphere, rather do (with constant potential temperature):
Equation of state with potential temperature:
\begin{equation}
  \label{eq:eos}
\Qrho(z) = \frac{p_{0}^{\frac{c_{p} - c_{v}}{c_{p}}} p^{\frac{c_{v}}{c_{p}}}{\left (z \right )}}{R \theta}
\end{equation}
Hydrostatic equilibrium with potential temperature:
\begin{equation}
  \label{eq:hydrostatic-balance-potT}
  \frac{d}{d z} p{\left (z \right )} = - \frac{g p_{0}^{\frac{R}{c_{p}}} p^{\frac{1}{\gamma}}{\left (z \right )}}{R \theta}
\end{equation}
Solution for pressure with initial condition of $p(0) = p_0$:
% \begin{equation*}
% p = \left(\frac{R \gamma \theta \left(\gamma^{\frac{\gamma}{\gamma - 1}} p_{0}\right)^{\frac{\gamma - 1}{\gamma}} - R \theta \left(\gamma^{\frac{\gamma}{\gamma - 1}} p_{0}\right)^{\frac{\gamma - 1}{\gamma}} - g \gamma p_{0}^{\frac{R}{c_{p}}} z \left(\gamma - 1\right) + g p_{0}^{\frac{R}{c_{p}}} z \left(\gamma - 1\right)}{R \gamma \theta \left(\gamma - 1\right)}\right)^{\frac{\gamma}{\gamma - 1}}
% \end{equation*}
\begin{equation}
p(z) =  \left(C - \frac{C}{\gamma} - \frac{g p_{0}^{\frac{R}{c_{p}}} z}{R \theta} + \frac{g p_{0}^{\frac{R}{c_{p}}} z}{R \gamma \theta}\right)^{\frac{\gamma}{\gamma - 1}},
\end{equation}
with constant of integration
\begin{equation}
  \label{eq:hydrostatic-balance-constant}
 C = \frac{\left(\gamma^{\frac{\gamma}{\gamma - 1}} p_{0}\right)^{\frac{\gamma - 1}{\gamma}}}{\gamma - 1},
\end{equation}
which can be obtained by setting $p(0) = p_0$.

>>>>

% TODO: Extract to settings.tex
\newcommand{\Qrho}{\rho}
\newcommand{\Qj}{\rho \bm{v}}
%\newcommand{\Qv}{\ensuremath{\Qrho^{-1} \Qj}}
\newcommand{\Qv}{\bm{v}}
\newcommand{\QE}{\rho E}
\newcommand{\potT}{\theta}
\newcommand{\backgroundPotT}{\overline{\theta}}
\newcommand{\pertubationPotT}{\theta'}
\newcommand{\stressT}{\bm{\sigma}}
\newcommand{\pressure}{p}
\newcommand{\maxConvEigen}{\vert \lambda_c^{\text{max}} \vert}
\newcommand{\maxViscEigen}{\vert \lambda_v^{\text{max}} \vert}

@@

We use the equation of state of an ideal gas
\begin{equation}
  \label{eq:eos}
  \pressure = (\gamma - 1) \left(\QE - 0.5 \left(\Qv \cdot \Qj \right) \right).
\end{equation}

The temperature $T$ relates to pressure and density by the ideal gas law
\begin{equation}
  \label{eq:temperature}
 \frac{\pressure}{\Qrho} = RT,
\end{equation}
where $R$ is the specific gas constant.

@@

\section{Scenarios}\label{sec:scenarios}

\subsection{Clouds}
The scenario is described in terms of potential temperature $\potT$
\begin{equation}
  \potT = T \frac{p_0}{p}^{R/c_p},
\end{equation}
where $p_o = \SI{10e5}{\Pa}$ is the reference pressure.
Solving for $T$ leads to
\begin{equation}
  \label{eq:potTToT}
  T = \potT (\frac{p}{p_0})^{R/c_p}.
\end{equation}
To allow for an easier description of the initial conditions we split the $\potT$ into a background state $\backgroundPotT$ and a pertubation $\pertubationPotT$
\begin{equation}
  \label{eq:potT-split}
  \potT = \backgroundPotT + \pertubationPotT.
\end{equation}

The initial conditions for pressure and density can be computed directly from the \textsc{pde} once few assumptions have been given.
We assume that $\backgroundPotT$ is constant over the entire domain.
Inserting the definition of potential temperature, given by \cref{eq:potTToT}, into the equation of state (see \cref{eq:eos}) leads us to
\begin{equation}
  \label{eq:eos-potT}
\Qrho(z) = \frac{p_{0}^{\frac{R}{c_{p}}} p^{\frac{1}{\gamma}}{\left (z \right )}}{R \backgroundPotT}.
\end{equation}
Flow in hydrostatic equilibrium can now be described by the ordinary differential equation(\textsc{ode})
\begin{align}
  \label{eq:hydrostatic-balance-potT}
  \frac{d}{d z} p{\left (z \right )} &= -g \Qrho(z)
                                     = 
        - \frac{g p_{0}^{\frac{R}{c_{p}}} p^{\frac{1}{\gamma}}{\left (z \right )}}{R \backgroundPotT}\\
  p{\left (0 \right )} &= p_0
\end{align}
where we use the reference pressure $p_0$ as the pressure on ground level.
This \textsc{ode} can be now simply solved by separation of variables.
After simplifying, we arrive at the initial condition for pressure
\begin{equation}
p(z) = \left(\left(1 - \frac{1}{\gamma}\right) \left(C - \frac{g p_{0}^{\frac{R}{c_{p}}} z}{R \backgroundPotT}\right)\right)^{\frac{c_{p}}{R}},
\end{equation}
with a constant of integration
\begin{equation}
  \label{eq:hydrostatic-balance-constant}
 C = \frac{c_{p} p_{0}^{\frac{R}{c_{p}}}}{R}.
\end{equation}

We now compute the pertubated potential temperature $\potT$ and use \cref{eq:potTToT} to convert it to temperature.
We then evaluate the density $\Qrho(\bm{x})$ with \cref{eq:temperature}.
The energy can be computed by inserting the previously computed pressure into the equation of state~(\ref{eq:eos}).
